---
title: 'Prediction Function'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# This file contains code that inputs all satellite data cleaned and sorted in
# python, and makes predictions from the selected model below. I prefer to use 
# the R statistical language to create the model and make predictions as I find 
# it more intuitive. Once the predictions are made, they are packaged into 
# another dataframe. This dataframe is then utilized in the python file 
# "Satellite regressions" to visualize all the predictions.


# Load required libraries:
```{r}

suppressMessages(library(caret))
suppressMessages(library(lme4))
suppressMessages(library(feather))
suppressMessages(library(arrow))
suppressMessages(library(dplyr))
suppressMessages(library(car))
suppressMessages(library(caret))
suppressMessages(library(MuMIn))

```




# Input the relevant data sets: Regional and global (Regional + Arctic)
```{r}

# Upload the data from a total data file:
# total_data <- read.csv("Total_Sat_Data_All_Var.csv")
total_data <- read.csv("MODEL_DATA.csv")
global_data <- read.csv("MODEL_DATA_GLOBAL.csv")

```


# Create a set of 'Prediction functions'. These functions will take large 
# data sets as an input, and return the predictions made by the model. The 
# model was validated and selected in the repository titled "Model_Selection":

# Regional Model
```{r}

Predict_fn_regional <- function(df) {
  
  # Select the relevant data for the model:
  regional_data <- total_data %>% select(TEP, Log_Chl, Temperature, Season, POC)
  
  # Create the model:
  regional_model <- lm(TEP ~ Log_Chl + Temperature + POC + Season, data = regional_data) 
  
  # Predictions:
  regional_predictions <- predict(regional_model, df, type = "response")  # Call the model and create predictions.
  regional_predictions <- as.data.frame(regional_predictions)  # Store the results in a data frame.
  
  # Extra data:
  regional_predictions$POC <- df$POC  # Add POC as a variable in the predictions data frame - for ratio derivations later on.
  regional_predictions$Temperature <- df$Temperature  # Add POC as a variable in the predictions data frame - for ratio derivations later on.
  regional_predictions$Log_Chl <- df$Log_Chl  # Add POC as a variable in the predictions data frame - for ratio derivations later on.
  
  return(regional_predictions)
  
}

```


# Global Model
```{r}
Predict_fn_global <- function(df) {
  
  # Select the relevant data for the model:
  global_data <- global_data %>% select(TEP, Season, Log_Chl, POC, Summer, Arctic, Area, Temperature)
  global_data <- na.omit(global_data)
  
  # Create the model:
  global_model <- lm(TEP ~ Log_Chl + POC + Temperature + Arctic, data = global_data)
  
  # Predictions:
  global_predictions <- predict(global_model, df, type = "response")  # Call the model and create predictions.
  global_predictions <- as.data.frame(global_predictions)  # Store the results in a data frame.
  
  # Extra data:
  global_predictions$POC <- df$POC  # Add POC as a variable in the predictions data frame - for ratio derivations later on.
  
  
  return(global_predictions)
  
}

```


# Prediction funciton application: Regional model
```{r}

# Upload all the satellite data from 2021 to be used as predictors:

# 8 day average spring 2021 regional:
df_Spring_8D_2021 <- arrow::read_feather('May_2021_LineP.feather')

# 8 day average summer 2021 regional:
df_Summer_8D_2021 <- arrow::read_feather('Aug_2021_LineP.feather')



# Place all the predictions into a list:
predictors <- vector(mode = "list", length = 2)
names(predictors) <- c("Spring_2021", "Summer_2021")

predictors[[1]] <- df_Spring_8D_2021; predictors[[2]] <- df_Summer_8D_2021



# Apply the function to the 2021 data, and convert the predictions to a feather file for transfer to python:
Regional_Predictions <-  lapply(predictors, Predict_fn_regional)

feather::write_feather(Regional_Predictions[[1]], 'May_2021_Predictions.feather')
feather::write_feather(Regional_Predictions[[2]], 'Aug_Sept_2021_Predictions.feather')

```


# Prediction funciton application: Global model
```{r}

# Upload all the satellite data from 2021 to be used as predictors:

# 8 day average spring 2021 global:
df_Spring_8D_2021 <- arrow::read_feather('May_2021_LineP.feather')

# 8 day average summer 2021 global:
df_Summer_8D_2021 <- arrow::read_feather('Aug_2021_LineP.feather')



# Place all the predictions into a list:
predictors <- vector(mode = "list", length = 2)
names(predictors) <- c("Spring_2021", "Summer_2021")

predictors[[1]] <- df_Summer_8D_2021; predictors[[2]] <- df_Summer_8D_2021



# Apply the function to the 2021 data, and convert the predictions to a feather file for transfer to python:
Global_Predictions <-  lapply(predictors, Predict_fn_global)

feather::write_feather(Global_Predictions[[1]], 'May_2021_Predictions_Global.feather')
feather::write_feather(Global_Predictions[[2]], 'Aug_Sept_2021_Predictions_Global.feather')

```


# Great! Now the functions have created large dataframes complete with target
# predictions within the .feather files. These files can now be read into 
# python for data visualization (this can also be done in R, but I prefer the 
# visualization techniques in python).
