---
title: "Stepwise_Regression_and_Model_Selection"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Upload required libraries:
```{r}

suppressMessages(library(ggplot2))
suppressMessages(library(caret))
suppressMessages(library(ggpubr))
suppressMessages(library(MuMIn))
suppressMessages(library(lme4))
suppressMessages(library(feather))
suppressMessages(library(arrow))
suppressMessages(library(dplyr))
suppressMessages(library(viridis))
suppressMessages(library(grid))
suppressMessages(library(car))

```


This code creates a series of both multivariate linear regression models and mixed effects models to predict the concentration of marine exopolymer gels important for both carbon sequestration in the ocean and the efficient functioning of desalination plants.

# Upload all the data, and sort into workable dataframes:
```{r}

# Upload the data from a total data file:
total_data <- read.csv("Total_Sat_Data_All_Var.csv")

UpperMixed_data <- total_data %>% select(TEP,  Log_Chl, Temperature, Nitrate, DO, Log_MLD, Season, Avg_PAR, Silicate, POC, Sigma)

```


# Step-wise Regressions models: Both forward and backward sequences for both types of models:
```{r}

# Define intercept-only model
intercept_only <- lm(TEP ~ 1, data = UpperMixed_data)  # linear
intercept_lmer <- lmer(TEP ~ 1 + (1|Season), data = UpperMixed_data)  # mixed effects

# Define model with all predictors
all <- lm(TEP ~ ., data = UpperMixed_data)  # linear
all_lmer <- lmer(TEP ~ . + (1|Season), data = UpperMixed_data)  # mixed effects

```


# Perform forward stepwise regression: Linear
```{r}
forward <- step(intercept_only, direction='forward', scope=formula(all), trace=0)
forward$anova
forward$coefficients

plot(forward$residuals ~ forward$fitted.values)

```


# Perform backward stepwise regression
```{r}

# Perform backward stepwise regression
backward <- step(all, direction ='backward', scope = formula(all), trace=0)
backward$anova
backward$coefficients

plot(backward$residuals ~ backward$fitted.values)

```


# Perform both backward and forward stepwise regression
```{r}

# Perform backward stepwise regression
both <- step(intercept_only, direction ='both', scope = formula(all), trace=0)
both$anova
both$coefficients

plot(both$residuals ~ both$fitted.values)

# This sequence comes to the same 'conclusions', or combinations of predictors, as the backwards regressions.
```


# Checking model assumptions: Equal variance and normal distribution of model residuals:
```{r}

# Get the linear model residuals:
UpperMixed_data$resid1 <- resid(forward)  
UpperMixed_data$predict1 <- predict(forward) 

UpperMixed_data$resid2 <- resid(backward)  
UpperMixed_data$predict2 <- predict(backward) 

# Residual, Q-Q, Cooks distance plots:
plot(forward)  # Residuals
plot(backward)

# Plot the distribution of the linear model residuals:
H_model1 <- ggplot(data = UpperMixed_data, aes(resid1)) + 
  geom_histogram(binwidth = 0.5, colour = "black", fill = "white") +
  xlab("Residuals") +
  ylab("Count") +
  theme_pubr()

H_model2 <- ggplot(data = UpperMixed_data, aes(resid2)) + 
  geom_histogram(binwidth = 0.5, colour = "black", fill = "white") +
  xlab("Residuals") +
  ylab("Count") +
  theme_pubr()

ggarrange(H_model1, H_model2)


# Use the Shapiro-Wilks test to test for normality distributions in the residuals:
shapiro.test(UpperMixed_data$resid1) 
shapiro.test(UpperMixed_data$resid2) 

summary(forward)
summary(backward)

# Check for autocorrelation in the predictors:
vif(forward)
vif(backward)
```

# Now we have selected the best possible combinations of variables to estimate marine exoplymers using linear regression. Here, I will introduce a series of mixed effects models based on the favoured variables above, and combinations of variables that are only available from remote sensing or satellite, in order to have a practical use for these models:

```{r}

# Satellite derived variables models:
Satellite_lmer <- lmer(TEP ~ Log_Chl + Temperature + POC + (1|Season), 
                    data = UpperMixed_data)  # Mixed effects

Satellite_lm<- lm(TEP ~ Log_Chl + Temperature + POC, 
                  data = UpperMixed_data) # Linear regression
summary(Satellite_lm)

# A simpler form of the satellite models:
Simple_Satellite_lmer <- lmer(TEP ~ Log_Chl + Temperature + (1|Season), 
                    data = UpperMixed_data)  # Mixed effects

Simple_Satellite_lm <- lm(TEP ~ Log_Chl + Temperature, 
                  data = UpperMixed_data)  # Linear regression

# Remote sensing models:
Remote_lmer <- lmer(TEP ~ Temperature + DO + Avg_PAR + Sigma + Log_MLD + (1|Season), 
               data = UpperMixed_data)  # Mixed effects

Remote_lm <- lm(TEP ~ Temperature + DO + Avg_PAR + Sigma + Log_MLD + factor(Season), 
               data = UpperMixed_data)  # Linear regression

# Selected forward regression model:
forward_lm <- lm(TEP ~ Log_Chl + Temperature + DO + POC + Avg_PAR + Nitrate +
                                  factor(Season), data = UpperMixed_data) # Mixed effects

forward_lmer <- lmer(TEP ~ Log_Chl + Temperature + DO + POC + Avg_PAR + Nitrate +
                                  (1|Season), data = UpperMixed_data)  # Linear regression

# Selected backward regression model:
backward_lm <- lm(TEP ~ Log_Chl + Temperature + DO + POC + Avg_PAR + Sigma +
                                   factor(Season), data = UpperMixed_data)  # Mixed effects

backward_lmer <- lmer(TEP ~ Log_Chl + Temperature + DO + POC + Avg_PAR + Sigma +
                                   (1|Season), data = UpperMixed_data)  # Linear regression


# Check the AIC and AICc values for each model:
AIC(Satellite_lmer, Satellite_lm, Simple_Satellite_lmer,  Simple_Satellite_lm, Remote_lmer, Remote_lm, 
    forward_lm, forward_lmer, backward_lm, backward_lmer)
AICc(Satellite_lmer, Satellite_lm, Simple_Satellite_lmer,  Simple_Satellite_lm, Remote_lmer, Remote_lm, 
    forward_lm, forward_lmer, backward_lm, backward_lmer)


# Check r squared for each model:
r.squaredGLMM(Satellite_lmer)
r.squaredGLMM(Satellite_lm)
r.squaredGLMM(Simple_Satellite_lmer)
r.squaredGLMM(Simple_Satellite_lm)
r.squaredGLMM(Remote_lmer)
r.squaredGLMM(Remote_lm)
r.squaredGLMM(forward_lm)
r.squaredGLMM(forward_lmer)
r.squaredGLMM(backward_lm)
r.squaredGLMM(backward_lmer)


# Test for normal distribution in residuals:

shapiro.test(resid(Satellite_lmer))
shapiro.test(resid(Satellite_lm))
shapiro.test(resid(Simple_Satellite_lmer))
shapiro.test(resid(Simple_Satellite_lm))
shapiro.test(resid(Remote_lmer))
shapiro.test(resid(Remote_lm))
shapiro.test(resid(forward_lm))
shapiro.test(resid(forward_lmer))
shapiro.test(resid(backward_lm))
shapiro.test(resid(backward_lmer))


# Check for multicolinearity among predictors:
car::vif(Satellite_lmer)
car::vif(Satellite_lm)
car::vif(Simple_Satellite_lmer)
car::vif(Simple_Satellite_lm)
car::vif(Remote_lmer)  
car::vif(Remote_lm)
car::vif(forward_lm)  
car::vif(forward_lmer)
car::vif(backward_lm)  
car::vif(backward_lmer)

# Examine model residuals and qqplots:
plot(Satellite_lmer)
plot(backward_lmer)

```

# the stepwise selection models clearly have some issues with multicollinearity among the predictor variables. Therefore, we should adress this issue by dropping the vairables of concern.

```{r}

backward_lm <- lm(TEP ~ Log_Chl + Temperature + DO + POC + Avg_PAR + Sigma, data = UpperMixed_data)  # Mixed effects
car::vif(backward_lm) 
r.squaredGLMM(backward_lm)

forward_lm <- lm(TEP ~ Log_Chl + Temperature + DO + POC + Avg_PAR + Nitrate, data = UpperMixed_data) # Mixed effects
car::vif(forward_lm) 
r.squaredGLMM(forward_lm)

# This improved the issues of mulitcollinearity. Now, lets run the AIC and AICc tests again:
AIC(Satellite_lmer, Satellite_lm, Simple_Satellite_lmer,  Simple_Satellite_lm, Remote_lmer, Remote_lm, 
    forward_lm, forward_lmer, backward_lm, backward_lmer)
AICc(Satellite_lmer, Satellite_lm, Simple_Satellite_lmer,  Simple_Satellite_lm, Remote_lmer, Remote_lm, 
    forward_lm, forward_lmer, backward_lm, backward_lmer)


```


# Now, I selected each of the most favourable models from each grouping (stepwise selection, satellite derived, remote sensing). Below, we will utilize a 
```{r}

# Containers for the predictions and accuracy measurements:

sample_train <- NULL

# Forward selections:
forward_selection_lm_predict <- NULL
forward_selection_lm_RMSE <- NULL
forward_selection_lm_MAE <- NULL

forward_selection_lmer_predict <- NULL
forward_selection_lmer_RMSE <- NULL
forward_selection_lmer_MAE <- NULL

# Backward selections:
backward_selection_lm_predict <- NULL
backward_selection_lm_RMSE <- NULL
backward_selection_lm_MAE <- NULL

backward_selection_lmer_predict <- NULL
backward_selection_lmer_RMSE <- NULL
backward_selection_lmer_MAE <- NULL

# Satellite derived:
sat_lm_predict <- NULL
sat_lm_RMSE <- NULL
sat_lm_MAE <- NULL

sat_lmer_predict <- NULL
sat_lmer_RMSE <- NULL
sat_lmer_MAE <- NULL

# Satellite derived: Basic
sat_lm_basic_predict <- NULL
sat_lm_basic_RMSE <- NULL
sat_lm_basic_MAE <- NULL

sat_lmer_basic_predict <- NULL
sat_lmer_basic_RMSE <- NULL
sat_lmer_basic_MAE <- NULL




# Loop through model prediction and cross validation 500x. The model is trained on 95% of the data, and validated on the other 5%, 500x.

for (i in 1:25) {
  
  #Creating a re-sampled data set from the total data:
  training.samples <- createDataPartition(UpperMixed_data$TEP, p = 0.95, list = FALSE)
  train.data1  <- UpperMixed_data[training.samples, ]  # Training set
  test.data1 <- UpperMixed_data[-training.samples, ]  # Testing set
  
  
  # Running the models and creating predictions, accuracy metrics on the partitioned training data:
  
  # Forward selections:
  forward_selection_lm<- lm(TEP ~ Log_Chl + Temperature + DO + POC + Avg_PAR + Nitrate, data = train.data1) 
  forward_selection_lmer<- lmer(TEP ~ Log_Chl + Temperature + DO + POC + Avg_PAR + Nitrate + (1|Season), data = train.data1) 
  
  # Backward selections:
  backward_selection_lm <- lm(TEP ~ Log_Chl + Temperature + DO + POC + Avg_PAR + Sigma, data = train.data1)  
  backward_selection_lmer <- lmer(TEP ~ Log_Chl + Temperature + DO + POC + Avg_PAR + Sigma + (1|Season), data = train.data1)  
  
  # Satellite derived: 
  sat_lm <- lm(TEP ~ Log_Chl + Temperature + POC, data = train.data1)  # Complex Satellite
  sat_lmer <- lmer(TEP ~ Log_Chl + Temperature + POC + (1|Season), data = train.data1)  # Complex Satellite
  
  # Satellite derived: basic
  sat_lm_basic <- lm(TEP ~ Log_Chl + Temperature, data = train.data1)  # Complex Satellite
  sat_lmer_basic <- lmer(TEP ~ Log_Chl + Temperature + (1|Season), data = train.data1)  # Complex Satellite

  
  
  # Model 'true' training points, applies to each model:
  sample_train <- c(sample_train, test.data1$TEP)
  
  
  
  # Model predictions and error calculations:
  
  # Forward selections: predictions and accuracy metrics
  forward_selection_lm_predict <- c(forward_selection_lm_predict, 
                                    predict(forward_selection_lm, test.data1, type = "response", allow.new.levels = TRUE))  # Predictions
  forward_selection_lm_RMSE <- c(forward_selection_lm_RMSE, RMSE(forward_selection_lm_predict, sample_train))  # RMSE:
  forward_selection_lm_MAE <- c(forward_selection_lm_MAE, MAE(forward_selection_lm_predict, sample_train))  # MAE:

  forward_selection_lmer_predict <- c(forward_selection_lmer_predict, 
                                    predict(forward_selection_lmer, test.data1, type = "response", allow.new.levels = TRUE))  # Predictions
  forward_selection_lmer_RMSE <- c(forward_selection_lmer_RMSE, RMSE(forward_selection_lmer_predict, sample_train))  # RMSE:
  forward_selection_lmer_MAE <- c(forward_selection_lmer_MAE, MAE(forward_selection_lmer_predict, sample_train))  # MAE:

  # Backward selections: predictions and accuracy metrics
  backward_selection_lm_predict <- c(backward_selection_lm_predict, 
                                    predict(backward_selection_lm, test.data1, type = "response", allow.new.levels = TRUE))  # Predictions
  backward_selection_lm_RMSE <- c(backward_selection_lm_RMSE, RMSE(backward_selection_lm_predict, sample_train))  # RMSE:
  backward_selection_lm_MAE <- c(backward_selection_lm_MAE, MAE(backward_selection_lm_predict, sample_train))  # MAE:

  backward_selection_lmer_predict <- c(backward_selection_lmer_predict, 
                                    predict(backward_selection_lmer, test.data1, type = "response", allow.new.levels = TRUE))  # Predictions
  backward_selection_lmer_RMSE <- c(backward_selection_lmer_RMSE, RMSE(backward_selection_lmer_predict, sample_train))  # RMSE:
  backward_selection_lmer_MAE <- c(backward_selection_lmer_MAE, MAE(backward_selection_lmer_predict, sample_train))  # MAE:

  # Satellite derived:
  sat_lm_predict <- c(sat_lm_predict, 
                                    predict(sat_lm, test.data1, type = "response", allow.new.levels = TRUE))  # Predictions
  sat_lm_RMSE <- c(sat_lm_RMSE, RMSE(sat_lm_predict, sample_train))  # RMSE:
  sat_lm_MAE <- c(sat_lm_MAE, MAE(sat_lm_predict, sample_train))  # MAE:

  sat_lmer_predict <- c(sat_lmer_predict, 
                                    predict(sat_lmer, test.data1, type = "response", allow.new.levels = TRUE))  # Predictions
  sat_lmer_RMSE <- c(sat_lmer_RMSE, RMSE(sat_lmer_predict, sample_train))  # RMSE:
  sat_lmer_MAE <- c(sat_lmer_MAE, MAE(sat_lmer_predict, sample_train))  # MAE:
  
  # Satellite derived: Basic:
  sat_lm_basic_predict <- c(sat_lm_basic_predict, 
                                    predict(sat_lm_basic, test.data1, type = "response", allow.new.levels = TRUE))  # Predictions
  sat_lm_basic_RMSE <- c(sat_lm_basic_RMSE, RMSE(sat_lm_basic_predict, sample_train))  # RMSE:
  sat_lm_basic_MAE <- c(sat_lm_basic_MAE, MAE(sat_lm_basic_predict, sample_train))  # MAE:

  sat_lmer_basic_predict <- c(sat_lmer_basic_predict, 
                                    predict(sat_lmer_basic, test.data1, type = "response", allow.new.levels = TRUE))  # Predictions
  sat_lmer_basic_RMSE <- c(sat_lmer_basic_RMSE, RMSE(sat_lmer_basic_predict, sample_train))  # RMSE:
  sat_lmer_basic_MAE <- c(sat_lmer_basic_MAE, MAE(sat_lmer_basic_predict, sample_train))  # MAE:
  
}

# The above loop generated 1000 different regression models with model coefficients, y intercepts and predictions, specified below for each of the models created: 
predictions <- cbind(forward_selection_lm_predict, forward_selection_lmer_predict, 
                     backward_selection_lm_predict, backward_selection_lmer_predict,
                     sat_lm_predict, sat_lmer_predict,
                     sat_lm_basic_predict, sat_lmer_basic_predict)
predictions <- as.data.frame(predictions)  # as a data frame

```


# Plot the results of the cross validation:
```{r}

forward_lm <- ggplot(predictions, aes(x = sample_train, y = forward_selection_lm_predict)) +
    geom_point(alpha = 0.6, colour = "black") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    ggtitle("1A: Chl (log) + Temperature + DO + POC + PAR + 
    Nitrate") +
    ylim(0, 200) +
    xlim(0, 200) +
    ylab("") +
    xlab("") +
    theme_pubr() +
    theme(plot.title=element_text(size=8,face="bold"))

forward_lmer <- ggplot(predictions, aes(x = sample_train, y = forward_selection_lmer_predict)) +
    geom_point(alpha = 1, colour = "black") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    ggtitle("1B: Chl (log) + Temperature + DO + POC + PAR + 
    Nitrate + (1|Season)") +
    ylim(0, 200) +
    xlim(0, 200) +
    ylab("") +
    xlab("") +
    theme_pubr() +
    theme(plot.title=element_text(size=8,face="bold"))



backward_lm <- ggplot(predictions, aes(x = sample_train, y = backward_selection_lm_predict)) +
    geom_point(alpha = 1, colour = "black") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    ggtitle("2A: Chl (log) + Temperature + DO + POC + PAR + 
    Density") +
    ylim(0, 200) +
    xlim(0, 200) +
    ylab("") +
    xlab("") +
    theme_pubr() +
    theme(plot.title=element_text(size=8,face="bold"))

backward_lmer <- ggplot(predictions, aes(x = sample_train, y = backward_selection_lmer_predict)) +
    geom_point(alpha = 0.6, colour = "black") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    ggtitle("(A) Chl + Temperature + DO + POC + PAR + Density + (1|Season)") +
    ylim(0, 200) +
    xlim(0, 200) +
    ylab("") +
    xlab("") +
    theme_pubr() +
    theme(plot.title=element_text(size=8,face="bold"))


sat_lm <- ggplot(predictions, aes(x = sample_train, y = sat_lm_predict)) +
    geom_point(alpha = 0.6, colour = "black") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    ggtitle("3A: Chl (log) + Temperature + POC") +
    ylim(0, 200) +
    xlim(0, 200) +
    ylab("") +
    xlab("") +
    theme_pubr() +
    theme(plot.title=element_text(size=8,face="bold"))

sat_lmer <- ggplot(predictions, aes(x = sample_train, y = sat_lmer_predict)) +
    geom_point(alpha = 0.6, colour = "black") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    ggtitle("(B) Chl + Temperature + POC + (1|Season)") +
    ylim(0, 200) +
    xlim(0, 200) +
    ylab("") +
    xlab("") +
    theme_pubr() +
    theme(plot.title=element_text(size=8,face="bold"))


sat_basic_lm <- ggplot(predictions, aes(x = sample_train, y = sat_lm_basic_predict)) +
    geom_point(alpha = 1, colour = "red") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    ggtitle("4A: Chl (log) + Temperature") +
    ylim(0, 200) +
    xlim(0, 200) +
    ylab("") +
    xlab("") +
    theme_pubr() +
    theme(plot.title=element_text(size=8,face="bold"))

sat_basic_lmer <- ggplot(predictions, aes(x = sample_train, y = sat_lmer_basic_predict)) +
    geom_point(alpha = 1, colour = "red") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    ggtitle("4B: Chl (log) + Temperature + POC + (1|Season)") +
    ylim(0, 200) +
    xlim(0, 200) +
    ylab("") +
    xlab("") +
    theme_pubr() +
    theme(plot.title=element_text(size=8,face="bold"))

```

# Plots:
```{r}
# Figure <- ggarrange(forward_lm, backward_lm, sat_lm, sat_basic_lm,
          # forward_lmer, backward_lmer, sat_basic_lmer, sat_basic_lmer, nrow = 4, ncol = 2)

Figure <- ggarrange(forward_lm, forward_lmer,
                    backward_lm, backward_lmer,
                    sat_lm, sat_lmer,
                    sat_basic_lm, sat_basic_lmer,
                     nrow = 4, ncol = 2)

Figure

pdf(file = "/Users/michaellivingston/Desktop/Cross_validations.pdf",   # The directory you want to save the file in
    width = 8, # The width of the plot in inches
    height = 12) # The height of the plot in inches

annotate_figure(Figure, 
                bottom = textGrob("Measured TEP ", gp = gpar(cex = 1)),
                left = textGrob("Predicted TEP ", gp = gpar(cex = 1), rot = 90))

dev.off()
```

# Plots:
```{r}

Figure2 <- ggarrange(backward_lmer, sat_lmer, nrow = 1, ncol = 2)

Figure2 <- annotate_figure(Figure2, 
                bottom = textGrob("Measured TEP ", gp = gpar(cex = 1)),
                left = textGrob("Predicted TEP ", gp = gpar(cex = 1), rot = 90))



pdf(file = "/Users/michaellivingston/Desktop/Cross_validations2.pdf",   # The directory you want to save the file in
    width = 10, # The width of the plot in inches
    height = 5) # The height of the plot in inches

Figure2


dev.off()

```


# Print the mean aboslute errors and root mean square errors for both models:
```{r}


# Create data frames for the accuracy measurements:
RMSE <- cbind(forward_selection_lm_RMSE, forward_selection_lmer_RMSE, 
              backward_selection_lm_RMSE, backward_selection_lmer_RMSE,
              sat_lm_RMSE, sat_lmer_RMSE,
              sat_lm_basic_RMSE, sat_lmer_basic_RMSE)
RMSE <- as.data.frame(RMSE)

MAE <- cbind(forward_selection_lm_MAE, forward_selection_lmer_MAE, 
              backward_selection_lm_MAE, backward_selection_lmer_MAE,
              sat_lm_MAE, sat_lmer_MAE,
              sat_lm_basic_MAE, sat_lmer_basic_MAE)
MAE <- as.data.frame(MAE)

mean_total <- mean(UpperMixed_data$TEP)  # Calculate mean of true values

# Get the average MAE and RMSE for each model:

MAE <- colMeans(MAE)
RMSE <- colMeans(RMSE)

print(MAE)
print(RMSE)

```




# Plot model residuals vs fitted values for each model:
```{r, fig.height=6, fig.width=12}

# Get the fitted and residual values:
UpperMixed_data$forward_resid <- resid(forward)
UpperMixed_data$forward_predict <- predict(forward)

forward_resid <- ggplot(UpperMixed_data, aes(x = forward_predict, y = forward_resid)) +
    geom_point(alpha = 1, colour = "blue") +
    geom_abline(slope = 0, intercept = 0, linetype = "dashed") +
    #ggtitle("A: Chl (log) + Nitrate + Temperature + DO + factor(Season)") +
    ylab("") +
    xlab("") +
    theme(plot.title = element_text(size=14, face="bold.italic")) +
    theme_pubr()

# Get the fitted and residual values:
UpperMixed_data$backward_resid <- resid(backward)
UpperMixed_data$backward_predict <- predict(backward)

backward_resid <- ggplot(UpperMixed_data, aes(x = backward_predict, y = backward_resid)) +
    geom_point(alpha = 1, colour = "blue") +
    geom_abline(slope = 0, intercept = 0, linetype = "dashed") +
    #ggtitle("B: Chl (log) + Temperature + DO + factor(Season)") +
    ylab("") +
    xlab("") +
    theme(plot.title = element_text(size=14, face="bold.italic")) +
    theme_pubr()

# Get the fitted and residual values:
UpperMixed_data$Theo3_resid <- resid(Theo_Model3)
UpperMixed_data$Theo3_predict <- predict(Theo_Model3)

Model3_resid <- ggplot(UpperMixed_data, aes(x = Theo3_predict, y = Theo3_resid)) +
    geom_point(alpha = 1, colour = "red") +
    geom_abline(slope = 0, intercept = 0, linetype = "dashed") +
    #ggtitle("C: Chl (log) + Temperature + DO + Nitrate + (1|Season)") +
    ylab("") +
    xlab("") +
    theme(plot.title = element_text(size=14, face="bold.italic")) +
    theme_pubr()

# Get the fitted and residual values:
UpperMixed_data$Theo4_resid <- resid(Theo_Model4)
UpperMixed_data$Theo4_predict <- predict(Theo_Model4)

Model4_resid <- ggplot(UpperMixed_data, aes(x = Theo4_predict, y = Theo4_resid)) +
    geom_point(alpha = 1, colour = "red") +
    geom_abline(slope = 0, intercept = 0, linetype = "dashed") +
    #ggtitle("D: Chl (log) + Temperature + (1|Season)") +
    ylab("") +
    xlab("") +
    theme(plot.title = element_text(size=14, face="bold.italic")) +
    theme_pubr()

# Get the fitted and residual values:
UpperMixed_data$Theo5_resid <- resid(Theo_Model5)
UpperMixed_data$Theo5_predict <- predict(Theo_Model5)

Model5_resid <- ggplot(UpperMixed_data, aes(x = Theo5_predict, y = Theo5_resid)) +
    geom_point(alpha = 1, colour = "black") +
    geom_abline(slope = 0, intercept = 0, linetype = "dashed") +
    #ggtitle("E: Chl (log)") + 
    ylab("") +
    xlab("") +
    theme(plot.title = element_text(size=14, face="bold.italic")) +
    theme_pubr()

# Get the fitted and residual values:
UpperMixed_data$Theo6_resid <- resid(Theo_Model6)
UpperMixed_data$Theo6_predict <- predict(Theo_Model6)

Model6_resid <- ggplot(UpperMixed_data, aes(x = Theo6_predict, y = Theo6_resid)) +
    geom_point(alpha = 1, colour = "black") +
    geom_abline(slope = 0, intercept = 0, linetype = "dashed") +
    #ggtitle("F: DO + Nitrate + factor(Season)") +
    ylab("") +
    xlab("") +
    theme(plot.title = element_text(size=14, face="bold.italic")) +
    theme_pubr()

Figure <- ggarrange(forward_resid, Model5_resid, Model3_resid, 
          backward_resid, Model6_resid, Model4_resid, nrow = 2, ncol = 3)


annotate_figure(Figure, 
                bottom = textGrob("Fitted Values ", gp = gpar(cex = 2)),
                left = textGrob("Residuals ", gp = gpar(cex = 2), rot = 90))

```

# Check for heteroscedasticity:
```{r}

car::ncvTest(forward)
car::ncvTest(backward)

lmtest::bptest(forward)
lmtest::bptest(backward)
lmtest::bptest(Theo_Model5)
lmtest::bptest(Theo_Model6)

```

# Get residual means:
```{r}

mean(resid(forward)) # mean
mean(resid(backward)) # mean
mean(resid(Theo_Model3)) # mean
mean(resid(Theo_Model4)) # mean
mean(resid(Theo_Model5)) # mean
mean(resid(Theo_Model6)) # mean

```


# Leave one out cross validation (LOOCV)
```{r}

# Specify the cross-validation method
ctrl <- trainControl(method = "LOOCV")

# Fit a regression model and use LOOCV to evaluate performance on the 'backwards' model:
backwards <- train(TEP ~ Log_Chl + Temperature + DO + POC + Avg_PAR + Sigma + Log_MLD + factor(Season), data = UpperMixed_data,
               method = "lm", trControl = ctrl)
print(backwards)
summary(backwards)


```


# Repeated k-fold cross validation:
```{r}

train_control <- trainControl(method ="repeatedcv", number = 10, repeats=30)

# train the model
backwards <- train(TEP ~ Log_Chl + Temperature + DO + POC + Avg_PAR + Sigma + Log_MLD + factor(Season), data = UpperMixed_data,
               method = "lm", trControl = train_control)
# summarize results
print(backwards)


```





